### 2019-10-10 old-introduction.tex
Initial Overleaf Import

So, how do we support more effective navigation of data
within spreadsheets?
One approach would be to try to integrate
an overview of the overall structure of the data along
with the spreadsheet~\cite{grudin2001partitioning}.
The resulting integration would then be a classical
overview+detail interface
where the spreadsheet is the detailed view.
Overview+detail interfaces are used
to facilitate navigation in various domains such as 
text editors and maps~\cite{cockburn2009review},
and enable users to manipulate
the overview or detailed view,
to perform high-level or low-level operations,
respectively.
Overview+detail interfaces
have been shown to be effective in these domains,
reducing cognitive load for users by
providing the big picture first,
helping them quickly assimilate the information space~\cite{cockburn2009review}.
\saj{In this paper, our goal is to integrate an overview plug-in with spreadsheets that captures the overall structure of the data while allowing interactions that minimize the difficulties caused by typical navigational operations like scrolling and steering.}

However, while an overview plug-in for spreadsheets
does seem appealing and natural,
developing it leads to several challenges.

\squishlist
\item {\em Overview modality.}
\saj{One approach for an overview is to simply add a
zoomed out version of the entire sheet as a pane on the side,
as in popular presentation software applications like Microsoft PowerPoint or text editors
like Sublime Text. 
Unfortunately, this approach
would not suffice for a spreadsheet---while it would help users make
somewhat more informed
scrolling decisions, the zoomed out presentation of the overview would be unreadable. An overview should provide a comprehensible big picture view of the spreadsheet.
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a global context of the user's current location
being displayed in the zoomed in detailed view~\cite{cockburn2009review}.
While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
view which allows objects to be represented differently at different scales. Since spreadsheets already display the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful. Rather, the spreadsheet overview should dynamically change as users sought more fine-grained or coarse-grained view of the overall structure of the data.}

\item {\em Construction of the overview.}
\saj{Given a spreadsheet
with many rows, 
one approach to constructing a dynamic overview is mapping rows of data to high level groups similar to online maps. In online maps, cities are grouped into states and states are grouped into countries forming a multi-granularity hierarchy. How do we automatically group spreadsheet rows together in a similar ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers? If the automatic grouping is not meaningful, how do we allow the users
to customize the overview
to construct meaningful groups? How do we facilitate interactions that enable users view the overview at multiple granularities?}

\item {\em Operations on the overview.}
\saj{Following the construction of a dynamic overview, the next challenge is to design simple interactions that achieve similar outcomes as scrolling and steering.
For example, an alternative to scrolling can be to leverage the groups of the overview to access the rows mapped to that group. How do we leverage the overview so that users can perform similar simple interaction to steer spreadsheet data and issue formula? As the granularity of the dynamic overview changes, how do we efficiently update the mapping from spreadsheet rows to the finer or coarser groups so that scrolling remains seamless? How do we recompute the results of a formula as the granularity changes without requiring the users to reissue the formula from scratch? How do we ensure such automatic recomputation scales with the size of the data?}

\item {\em Seamless integration as a plug-in.}
Finally, the design of the overview
should be generic so that it can be integrated
with any existing spreadsheet tool on any spreadsheet
dataset without impacting
existing functionalities or look-and-feel.
Any new interactions introduced should be consistent
with traditional spreadsheet semantics,
and complement existing spreadsheet interactions.
Interactions between the overview and the raw spreadsheet
should be coordinated to the extent possible.
\squishend

\begin{comment}
Since we want to allow users to control how the overview
is constructed, we can let them select the grouping
attribute(s). Following this, we need mechanisms to automatically group nearby rows
into meaningful bins.
This necessitates
{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
zooming (unlike text editors or presentation software),
on the overview view instead of the detailed view
(unlike map tools).
Several domains, \eg online maps and program visualization,
adopt semantic zooming
to ensure improved readability and
comprehension of the information displayed~\cite{summers2003experimental}.
However, to the best of our knowledge, no prior work
has ever developed semantic zooming for spreadsheet data.
Customizing these bins will also eliminate
visual discontinuities: users can simply arrange the bins
the way they want and keep them near each other
for easy comparison.
For this, we take inspiration from aggregation
in relational databases. In relational databases, each ``group'' is associated
with one or more aggregates. As long as the groups are
appropriately constructed (using the mechanisms alluded to above),
formulae can be issued once without requiring any steering, and
automatically maintained for all groups when
zooming in and zooming out. 
\end{comment}

\stitle{\noah.} We address the aforementioned challenges
in our tool \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
\saj{a multi-granularity overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a fine-grained or coarse grained perspective of the data distribution. The distribution at each granularity is captured by a histogram,
enabling users to assimilate the data. Each bin (group) of the histogram is mapped to a collection of rows in the spreadsheet.} Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface. Instead of steering to analyze the data,
users can issue formulae
on the overview with interactions similar to pivot table construction,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b). 
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

>>>>

So, how do we support more effective navigation of data
within spreadsheets?
One approach would be to try to integrate
an overview of the overall structure of the data along
with the spreadsheet~\cite{grudin2001partitioning}.
The resulting integration would then be a classical
overview+detail interface
where the spreadsheet is the detailed view.
Overview+detail interfaces are used
to facilitate navigation in various domains such as 
text editors and maps~\cite{cockburn2009review},
and enable users to manipulate
the overview or detailed view,
to perform high-level or low-level operations,
respectively.
Overview+detail interfaces
have been shown to be effective in these domains,
reducing cognitive load for users by
providing the big picture first,
helping them quickly assimilate the information space~\cite{cockburn2009review}.
In this paper, our goal is to extend
existing spreadsheets
with an overview plug-in, giving an overview+detail interface.

However, while an overview plug-in for spreadsheets
does seem appealing and natural,
developing it leads to several challenges.
\squishlist
\item {\em Overview modality.}
One approach for an overview is to simply add a
zoomed out version of the entire sheet as a pane on the side,
as in popular presentation software applications like Microsoft PowerPoint or text editors
like Sublime Text. 
Unfortunately, this approach
would not suffice for a spreadsheet---while it would help users make
somewhat more informed
scrolling decisions, the zoomed out presentation of the overview would be unreadable.
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a \saj{global} context of the user's current location
\saj{being displayed} in the \saj{zoomed in} detailed view~\cite{cockburn2009review}.
Users can perform zooming operations on the detailed
view as the overview remains static.
Since spreadsheets already d  isplay the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful.
Instead, the goal of our spreadsheet
overview is to provide a big picture view of the spreadsheet,
capturing the
data distribution and \saj{augmenting it with summarization},
while enabling meaningful interactions
for navigating the data.
This necessitates
{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
zooming (unlike text editors or presentation software),
on the overview view instead of the detailed view
(unlike map tools).
Several domains, \eg online maps and program visualization,
adopt semantic zooming
to ensure improved readability and
comprehension of the information displayed~\cite{summers2003experimental}.
However, to the best of our knowledge, no prior work
has ever developed semantic zooming for spreadsheet data.

\item {\em Construction of the overview.}
Given a spreadsheet
with many rows,
it is not clear how to employ semantic zooming
to construct the overview.
How do we group rows together in a ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers?
Since we want to allow users to control how the overview
is constructed, we can let them select the grouping
attribute(s).
Following this, we need mechanisms to automatically group nearby rows
into meaningful bins.
Further, the bins need to be constructed
at various granularities,
allowing users to zoom in and zoom out of the overview.
We also need to allow the users
to customize the bins
to construct meaningful categories, while providing
interpretable and intelligent defaults.
Customizing these bins will also eliminate
visual discontinuities: users can simply arrange the bins
the way they want and keep them near each other
for easy comparison.

\item {\em Operations on the overview.}
While the scrolling issues are partially addressed
by having the overview, the steering issues remain unaddressed.
How do we leverage the overview to allow users
to select regions of interest for issuing formulae
without steering
across multiple screens?
As the granularity of the overview changes (during zoom in/out)
how do we ensure that users don't have to reissue the formula from scratch?
For this, we take inspiration from aggregation
in relational databases. In relational databases, each ``group'' is associated
with one or more aggregates. As long as the groups are
appropriately constructed (using the mechanisms alluded to above),
formulae can be issued once without requiring any steering, and
automatically maintained for all groups when
zooming in and zooming out.

\item {\em Seamless integration as a plug-in.}
Finally, the design of the overview
should be generic so that it can be integrated
with any existing spreadsheet tool on any spreadsheet
dataset without impacting
existing functionalities or look-and-feel.
Any new interactions introduced should be consistent
with traditional spreadsheet semantics,
and complement existing spreadsheet interactions.
Interactions between the overview and the raw spreadsheet
should be coordinated to the extent possible.

\squishend

\noindent
We address the aforementioned challenges
in our tool \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
an overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a).
The overview is a histogram
and depicts the overall data distribution
for the chosen attribute,
enabling users to assimilate the data.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface.
% \noah leverages {\scshape DataSpread}'s
% underlying spatial mapping of data to ensure coordination between the overview and the spreadsheet so that users can relate the interactions on the overview with the raw data.
Instead of steering to analyze the data,
users can issue aggregation operations
on the overview,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b).
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

### 2019-10-10 system.tex
Initial Overleaf Import

In this supplementary material we highlight the use cases that \noah
is suited for, the in-situ chart representations within
the overview, \noah’s underlying system architecture,
and statistical significance results from our user study.
\section*{\noah Use Cases}
As mentioned in Section~\ref{sec:design}, the goal of designing 
an overview interface is to 
not only enable seamless navigation 
but also to facilitate functionalities to help users visually seek information about the data. Therefore, we draw parallels to the typology of abstract visualization tasks~\cite{brehmer2013multi} to identify the use cases that \noah is best suited for. 
We are specifically interested in exploring the different task purposes mentioned in the typology. The task purposes can be specified at three different levels: consume and produce (high-level), search (mid-level), and query (low-level). Table~\ref{tab:scope} captures all three levels and highlights where \noah can be used.

@@

\section*{Explaining the Chart Representations}

@@

\section*{Implementation and  Architecture}
\label{sec:sys}
In this section, we provide an overview of the infrastructure of \noah. 
We integrate \noah as a data exploration plugin within{\scshape DataSpread}~\cite{dataspread}, 
an open-source scalable web-based spreadsheet.

@@

\subsection*{System Architecture}

@@

\section*{Results}
In this section, we evaluate the statistical significance of the task performance results and survey responses presented in Section~\ref{sec:results}. 
\begin{figure}
    \centering
    %\vspace{-10pt}
    \includegraphics[trim=0 57 0 145,clip,width=\linewidth]{images/userinfo-ip-1.pdf}
   \caption{Intra-participant submission time differences between \noah and Excel across the quiz tasks.}
   \label{fig:ipdiff}
 \end{figure}

\subsection{Intra-participant differences.}
Figure~\ref{fig:ipdiff} depicts the intra-participant submission time differences between \noah and Excel, across all the quiz tasks.
For a given task and a participant, the corresponding circle denotes by what percentage (between $0.30\%$ to $94.33\%$) a system is faster than the competing system. The color of the circle denotes which system was faster (green: \noah faster, yellow: Excel faster). The larger the circle the faster the corresponding system is. For example, participant $P9$'s submission time was $91.78\%$ faster with \noah for the \cmpB task and $72.62\%$ faster with Excel for the \cmpA task. Figure~\ref{fig:ipdiff} confirms that majority of the submission times using \noah were faster compared to Excel---nineteen out of the twenty participants completed at least four tasks in less time using \noah compared to Excel. The submission time difference is even more apparent for the steer, identify, and \cmpB tasks.

For \cmpA, customize, and the second steer task, less than a third of the participants' submission times were faster using Excel. 
For the second steer tasks, four out of the six participants that submitted answers quickly using Excel compared to \noah, utilized the \emph{autosum} shortcut for the birdstrikes dataset. For the \cmpA tasks, a number of participants ($N=3$) that required more time to submit answer using \noah, first used the bin customization feature to view all the unique bins which contributed to the higher submission time compared to Excel. On the other hand, as explained earlier, the unfamiliarity with the bin customization feature contributed to higher submission time using \noah for six participants (yellow circles).

\subsection{Statistical Significance: Task Performance}
Since we conducted our user study on a small population (20 participants), we further evaluated the statistical significance of the task performance results, \ie accuracy and completion time. 
To measure the significance of the task completion times, we ran \emph{Mann-Whitney's U test} (as completion times did not follow a normal distribution). \emph{For all of the tasks except the customize task, we found a significant effect of the tools, \ie the response times for the tasks significantly differed by the choice of the tool} (see Table~\ref{tab:statTest}). 
On the other hand, we ran the \emph{Fisher's exact test} that measures the statistical significance of categorical data (0/1 accuracy): \emph{only for steer tasks, the percentage of the accuracy of submissions significantly differed by the choice of the tool}.

>>>>

In this supplementary material we highlight the use cases that \noah
is suited for, the in-situ chart representations within
the overview, \noah’s underlying system architecture,
and statistical significance results from our user study.
\section{\noah Use Cases}
As mentioned in Section~\ref{sec:design}, the goal of designing 
an overview interface is to 
not only enable seamless navigation 
but also to facilitate functionalities to help users visually seek information about the data. Therefore, we draw parallels to the typology of abstract visualization tasks~\cite{brehmer2013multi} to identify the use cases that \noah is best suited for. 
We are specifically interested in exploring the different task purposes mentioned in the typology. The task purposes can be specified at three different levels: consume and produce (high-level), search (mid-level), and query (low-level). Table~\ref{tab:scope} captures all three levels and highlights where \noah can be used.

\begin{table}[!htb]{\scriptsize}
\caption{Uses case of \noah.}
\label{tab:scope}
\centering
\begin{tabular}{c c}
\hline
Purpose & Use Cases   \\ \hline
\emph{Consume}         & \code{discover} (\checkmark), \code{present} (\checkmark), \code{enjoy} (\checkmark)\\
\emph{Produce}         & \code{create} (\checkmark), \code{annotate} ($\times$)\\
\emph{Search}            & \code{browse} (\checkmark), \code{explore (\checkmark)}, \code{locate} ($\times$), \code{lookup} ($\times$)\\
\emph{Query}    &  \code{identify} (\checkmark), \code{summarize} (\checkmark), \code{compare} (\checkmark)  \\ \hline
\end{tabular}
\end{table}

@@

\section{Explaining the Chart Representations}

@@

\section{Implementation and  Architecture}
\label{sec:sys}
In this section, we provide an overview of the infrastructure of \noah. 
We integrate \noah as a data exploration plugin within{\scshape DataSpread}~\cite{dataspread}, 
an open-source scalable web-based spreadsheet.

@@

\subsection{System Architecture}

@@

\section{Results}
In this section, we evaluate the statistical significance of the task performance results and survey responses presented in Section~\ref{sec:results}. 
\subsection{Statistical Significance: Task Performance}
Since we conducted our user study on a small population (20 participants), we further evaluated the statistical significance of the task performance results, \ie accuracy and completion time. 
To measure the significance of the task completion times, we ran \emph{Mann-Whitney's U test} (as completion times did not follow a normal distribution). \emph{For all of the tasks except the customize task, we found a significant effect of the tools, \ie the response times for the tasks significantly differed by the choice of the tool} (see Table~\ref{tab:statTest}). 
On the other hand, we ran the \emph{Fisher's exact test} that measures the statistical significance of categorical data (0/1 accuracy): \emph{only for steer tasks, the percentage of the accuracy of submissions significantly differed by the choice of the tool}.
